### 2014-08-06 docs/paper/text/introduction.tex
Checkpoint. Working on writing an introduction.

\section{Introduction}

Query workload and disk layout have a significant impact on database
performance.  For table-based, relational databases, this fact has led database
designers to develop alternative approaches for storage layout: row-oriented
storage is more efficient when queries access many attributes from a small
number of records, and column-oriented storage is more efficient when queries
access a small number of attributes from many records.

Recently, there has been increased interest in graph databases in general, and
\emph{interaction graph} databases in particular.  An \emph{interaction graph}
is an append-only graph, where new edges and vertices are added as time
progresses.  Thus, interaction graph encode the relationships over time between
entities in their structure, making them ideal for use in social networking,
telecommunications, or modeling the world-wide web.

Unfortunately, although interaction graph databases, like relational databases,
are the target of diverse query workloads, there is no clear correspondence to a
row-oriented or column-oriented storage layout.
Typical interaction graphs have data (i.e., \emph{attributes} ) associated with
their vertices and edges. Most algorithms access this data as they traverse the
graph.

In an interaction graph database organized as blocks of temporal neighbor lists
(as in Gedik et al.~\cite{gedik14}), it is important to store the edge and
vertex attributes locally. For  instance, if the edge properties were kept separately
in a relational table, there will be almost no benefit to the locality
optimizations performed for block organization (discussed in Gedik et
al.~\cite{gedik14}, as we would have to go back and forth between the disk
blocks to access the edge attributes).%

On the other hand, putting all the edge attributes into the disk blocks
containing the graph structure is expected to cause significant overhead when
only a few of these attributes are read. This is somewhat similar to the

% However, not all of the data is
% accessed by all the queries. Typically, there are correlations among the
% attributes accessed by different queries, such as Q1 and Q5 accessing
% attributes a and b, and Q2, Q3, Q4 accessing attributes c and d, and so on.%

Typically, a graph database stores entities as vertices,
relationships as edges, and additional information in labels that may be
associated with either vertices or edges.

They store entities as vertices,
relationships as edges, and additional information in labels that may be
associated with either vertices or edges.

An \emph{interaction graph} is an append-only graph, where new edges and
vertices are added as time progresses.%

Typical interaction graphs have data associated with their vertices and edges.
We call this data \emph{attributes} or \emph{properties}. Most algorithms
access this data as they traverse the graph. However, not all of the data is
accessed by all the queries. Typically, there are correlations among the
attributes accessed by different queries, such as Q1 and Q5 accessing
attributes a and b, and Q2, Q3, Q4 accessing attributes c and d, and so on.%

% In an interaction graph database organized as blocks of temporal neighbor lists
% (as in Gedik et al.~\cite{gedik14}), it is important to store the edge and
% vertex properties locally. For  instance, if the edge properties are kept away
% in a relational table, there will be almost no benefit to the locality
% optimizations performed for block organization (discussed in Gedik et
% al.~\cite{gedik14}, as we would have to go back and forth between the disk
% blocks to access the edge attributes).%

On the other hand, putting all the edge attributes into the disk blocks
containing the graph structure is expected to cause significant overhead when
only a few of these attributes are read. This is somewhat similar to the 
problem with row-oriented databases, where the entire row needs to be accessed
from the disk despite the query needing only a small fraction of it.
Unfortunately, there is no clear correspondence to a column-oriented database
layout for the graph databases.%

An important characteristic of interaction graphs are that they are temporal.
As such, the co-access correlations for the attributes can be different for
different temporal regions. It might also be unknown at the insertion time and
may be discovered later depending on the workload. This points to the need for
adaptively optimizing the layout (somewhat similar to H2O \cite{alagiannis14}
and HYRISE~\cite{grund10}).%

I imagine a solution to this problem, which I name the `rail layout'. The idea
is to start with large blocks that contain the entire data. As we learn about
the access properties for different time regions, we might start splitting
such blocks into smaller blocks that run parallel to each other, almost like
having two or more graph databases for certain time regions, each containing a
 different subset of the attributes, but with a link between them in case a
query needs to access both.