### 2014-08-07 docs/workshop/text/abstract.tex
Renamed cidr to workshop

This paper presents an algorithm for adaptively optimizing disk block storage for interaction graphs.  An interaction graph is an append-only, temporal graph structure that encodes the relationships between entities over time. The algorithm makes use of a novel storage scheme, called a rail layout, that partitions disk blocks into smaller blocks that contain the graph structure, but only a subset of attributes. In the ideal case, the database need only read the attributes relevant to answering a particular query. The paper presents an integer linear program formulation to determine an optimal partitioning of attributes for a given query workload. Simulations show that the rail layout scheme significantly reduces latency due to query I/O for interaction graphs.