### 2014-08-14 docs/paper/text/evaluation.tex
merged the text from the workshop. fixed bibtex bugs

\section{Evaluation}

\subsection{I/O comparison}
\begin{alltt}\scriptsize
optimal vs single partition vs. attribute per partition
Compute the I/O for different workloads,
3 values averaged at the end
Also compute the storage...
\end{alltt}

\subsection{LP Solver scalability}
\begin{alltt}\scriptsize
# of attrbutes vs solving time:
\end{alltt}