### 2014-09-11 docs/edbt2015/text/heuristic.tex
paper skeleton

\section{Heuristic Solution}\label{sec:heuristic}