### 2014-09-11 docs/edbt2015/text/paper.tex
paper skeleton

\input{config}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Railway: Adaptive Disk Storage for Interaction Graphs}

\numberofauthors{2} 
\author{
\alignauthor
Robert Soul\'{e}\\
       \affaddr{University of Lugano}\\
       \email{robert.soule@usi.ch}  
\alignauthor
Bu\u{g}ra Gedik\\
       \affaddr{Bilkent University}\\
       \email{bgedik@cs.bilkent.edu.tr}
}

\maketitle
\begin{abstract}
  \input{abstract}            
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\input{introduction}

\begin{alltt}\scriptsize
- Intro
    - What is an interaction graphs
    - Relationship graphs vs interaction graphs 
        - interactions get big
\end{alltt}

\input{system.tex}
\input{problem.tex}
\input{optimal.tex}
\input{heuristic.tex}
\input{evaluation.tex}
\input{related.tex}
\input{conclusion.tex}

\bibliographystyle{abbrv}
\bibliography{paper}  
\end{document}