### 2014-09-11 docs/edbt2015/text/related.tex

\section{Related Work}\label{sec:related}

\begin{alltt}\scriptsize
Gedik et al.~\cite{gedik14}
H2O \cite{alagiannis14}
HYRISE~\cite{grund10}
Bornea et al.~\cite{bornea13}
\end{alltt}