### 2014-09-16 docs/edbt2015/text/optimal.tex
renamed sections

\section{Optimal Railway Design}\label{sec:optimal}

\begin{alltt}\scriptsize
- Layout or partitioning problem
    - Cost model
\end{alltt}