### 2014-09-16 docs/edbt2015/text/system.tex
outline of the systems section

\section{System Overview}\label{sec:system}
\begin{alltt}\scriptsize
- System overview
    - How interaction graph are currently stored
    - Introduce the rail layout
    - Add the Figure~\ref{fig:}
    - Query model
\end{alltt}