### 2014-09-20 docs/edbt2015/text/heuristic.tex
started on heuristic solution

\section{Heuristic Solution}\label{sec:heuristic}

\begin{alltt}
- Talk about the scalability of the optimal solution
- Do we have a running time?
- Introduce a heuristic approximation approach
\end{alltt}

\subsection{Non-Overlapping Partitions}\label{subsec:nov-heuristic}
\noindent

\paragraph*{Approach$\,$}
TODO

\begin{algorithm}[ht]
\scriptsize
\caption{Algorithm for partitioning blocks into sub-blocks with non-overlapping attributes.}
\label{alg:non-overlappingP}
\KwData{$B$: block, $Q$: set of queries}
$c^*\leftarrow \infty$ \tcp*{Lowest cost over all \# of partitions}
\For(\tcp*[f]{For each possible \# of partitions}){$k=1$ to $|A|$}{
   $R[i]\leftarrow \emptyset, \forall i\in [1..k]$ \tcp*{Initialize partitions}
   \For(\tcp*[f]{For each attribute}){$a \in A$\textnormal{, in decr.\/ order of }$f(a)$}{
      $c\leftarrow \infty$ \tcp*{Lowest cost over all assignments}  
      $j\leftarrow -1$ \tcp*{Best partition assignment}
      \For(\tcp*[f]{For each partition assignment}){$i\in [1..k]$} {
         $R[i]\leftarrow R[i] \cup \{a\}$\tcp*{Assign attribute}
         \If(\tcp*[f]{If query cost is lower}){$L(R, B, Q)<c$}{
            $c\leftarrow L(R, B, Q)$\tcp*{Update the lowest cost}
            $j\leftarrow i$\tcp*{Update the best partition}
         }
         $R[i]\leftarrow R[i] \setminus \{a\}$\tcp*{Un-assign attribute}
      }
      $R[j]\leftarrow R[j] \cup \{a\}$\tcp*{Assign to best partition}
   }
   \lIf(\tcp*[f]{If solution infeasible}){$H(R, B, Q)>\alpha$}{\textbf{break}}
   \If(\tcp*[f]{If solution has lower cost}){$L(R, B, Q)<c^*$}{
     $c^* \leftarrow L(R, B, Q)$\tcp*{Update the lowest cost}
     $\mathcal{P}(B)\leftarrow R$\tcp*{Update the best partitioning}
   } 
}
\Return $\mathcal{P}(B)$ \tcp*{Final set of sub-blocks}
\end{algorithm}

\subsection{Overlapping Partitions}\label{subsec:ov-heuristic}
\noindent

\paragraph*{Approach$\,$} We start the algorithm with a   partitioning based
on what queries we have seen. Every query gets its own sub-block. This is the
``ideal'' partitioning, because the I/O cost would be minimized for every
query that we would have seen. The algorithm iteratively combines the two
partitions that are closest together.  After each combination of partitions,
the algorithm calculate the storage overhead for the partitioning. The
algorithm stops when the  storage cost is below some specified threshold.  The
result is the block partitioning.

\begin{algorithm}[ht]
\scriptsize
\caption{Algorithm for partitioning blocks into sub-blocks with overlapping attributes.}
\label{alg:overlappingP}
\KwData{$B$: block, $Q$: set of queries}
$\mathcal{P}(B) \leftarrow \{q.A: q\in Q\}$ \tcp*{Each query gets its own sub-block}
\While(\tcp*[f]{Until storage overhead is below $\alpha$}){$H(\mathcal{P},B) > \alpha$}{
  $c^{*}\gets \infty $ \tcp*{Lowest cost over all sub-block pairs}
  $(b_x,b_y)\gets (\emptyset,\emptyset)$ \tcp*{Sub-block pair with the lowest cost}
  \For(\tcp*[f]{For each pair of blocks}){$\{b_i,b_j\}\in\mathcal{P}(B)$}{
    $\mathcal{P'}(B) \leftarrow \mathcal{P}(B) \setminus \{b_i, b_j\} \cup \{b_i \cup b_j\}$\\
    $c\gets \frac{L(\mathcal{P}',B,Q)-L(\mathcal{P},B,Q)}{H(\mathcal{P},B)-H(\mathcal{P}',B)}$\tcp*{Cost of merge}
    \If(\tcp*[f]{Cost is lower}){$c<c^{*}$}{
        $c^{*}\gets c$\tcp*{Update the lowest cost}
        $(b_x,b_y)\gets (b_i,b_j)$\tcp*{Update the best pair}
    }
  }
  $\mathcal{P}(B) \leftarrow \mathcal{P}(B) \setminus \{b_x, b_y\} \cup \{b_x \cup b_y\}$ \\
}
\Return $ \mathcal{P}(B)$  \tcp*{Final set of sub-blocks}
\end{algorithm}