### 2014-09-21 docs/edbt2015/text/related.tex
Added some related work text. did some bibtex cleanup

\section{Related Work}\label{sec:related}

The railway layout and algorithms build on our prior work~\cite{gedik14}, which
organized the disk layout for an interaction graph databases to improve access
locality. The railway layout extends the previous design, by enabling the system
to adapt the layout for changing workloads.

Our adaptation scheme is similar to work on adaptive layouts for relational
database. The $H_2O$~\cite{alagiannis14} system can adapt its data layout into
three types, row-major, column-major, or groups of columns, depending on the
workload. HYRISE~\cite{grund10} provides a similar adaptive layout scheme for an
in-memory relational database. Both systems use heuristic, iterative solutions
to determine partitioning. The railway layout scheme differs, in that it
targets interaction graphs, and we present optimal solutions, in addition to 
heuristic solutions.

 \begin{alltt}\scriptsize
% Gedik et al.~\cite{gedik14}
% H2O \cite{alagiannis14}
% HYRISE~\cite{grund10}
 Bornea et al.~\cite{bornea13}
 \end{alltt}